The role of continuous developmental change in explaining discontinuities in word learning

### 2019-01-29 cogsci/cogsci.tex
second draft cogsci

\title{Continuous Development in Learning Similar-sounding Words}

\usepackage{tipa}

@@

\begin{abstract}
Stager and Werker (1997) first showed that children's early ability to
distinguish similar sounds do not automatically translate into word
learning skills. Early explanations proposed that children may not be
able to encode subtle phonetic contrasts when learning novel word
meanings, thus suggesting a discontinuous/stage-like pattern of
development. However, later work has revealed (e.g., through using
simpler testing methods) that children do encode such contrasts, thus
favoring a rather a continuous pattern of development. The present study
proposes a precise probabilistic model describing how development may
proceed in a continuous fashion across the lifespan. The purpose of the
model is to account for previously documented facts while providing new
predictions. We collected data from both preschool children and adults,
and we showed that the model can explain various patterns of learning
both within the same age and across development. Our work highlights the
role of computational modeling in advancing our understanding of
development both through organizing exsinting knowledge and through
generating new principled hypotheses.

@@

Over the first year of life, children become sensitive to the phonetic
variations that are used to distinguish meanings in their native
language (Werker \& Tees, 1984). One could imagine that these perceptual
skills would be automatically applied to the task of word learning.
However, developmental data show that 14 m.o children find it
challenging to associate minimally different (but perceptually
discriminated) sounds such as ``bin'' and ``din'' to different objects
(Stager \& Werker, 1997).

Several factors can explain this finding. For example it is possible
that the task of meaning learning increases cognitive demands on
children (compared to a simple perceptual discrimination). In
particular, it requires paying attention to both the sounds and the
corresponding objects, which may hinder precise encoding in memory of
some phonetic details (Hofer \& Levy, 2017; Stager \& Werker, 1997).
Additional difficulty might arise from ambiguous phonological boundaries
at this stage of development (e.g., Rost \& McMurray, 2009), or from
uncertainty about the referential status of the novel word (Fennell \&
Waxman, 2010).

Regardless of the exact explanation, it is generally accepted that by
around 17 m.o, children succeed in the same task (Werker, Fennell,
Corcoran, \& Stager, 2002). What could be the mechanism of development?
Early accounts assumed--whether implicitly or explicitly-- that children
encode words in a binary way: they either fail or succeed in encoding
the relevant phonetic details (simultaneously with the meanings). This
account suggested a discontinuous/stage-like pattern of development
whereby younger children fail to encode the contrastive phonetic detail,
whereas older children succeed.

Nevertheless, subsequent findings suggest otherwise. On the one hand,
14-month-olds--who typically fail in the original task--succeed when an
easier \emph{testing} method is used, even under the same
\emph{learning} conditions (Yoshida, Fennell, Swingley, \& Werker,
2009). They also succeed when uncertainty is mitigated via
disambiguating cues (e.g., Thiessen, 2007). On the other hand, adults
show patterns of learning similar to those shown by 14-month-olds when
the task is more challenging and when similarity between words increases
(Pajak, Creel, \& Levy, 2016; White, Yee, Blumstein, \& Morgan, 2013).

These evidence point towards another scenario, where the representations
are encoded in a probabilistic (rather than binary) way, and where
development is continuous, rather than stage-like (Pajak et al., 2016;
Swingley, 2007; Yoshida et al., 2009). On this account, correct
representations are learned early in development, but these
representations are encoded with higher uncertainty in younger children,
leading to apparent failure in relatively demanding tasks. Development
is a continuous process whereby the initial noisy representations become
more precise. Crucially, more precise representation are still
imperfect: Even with adults show low accuracy learning when the sounds
are subtle enough, e.g., non-native sounds (Pajak et al., 2016).

We provide an intuitive illustration of how such an account explains
patterns of learning and development in Figure \ref{fig:illus}. We
observe low accuracy in word learning when the perceptual distance
between the labels is small relative to the uncertainty/tolerance with
which these labels are encoded. For example, in Stager and Werker's
original experiment, children are supposed to associate label 1
(``bih'') and label 2 (``dih'') with object 1 and object 2,
respectively. Though children could learn that the label ``bih'' is a
better match to object 1 than ``dih'', they could still judge the sound
``dih'' as a plausible instance of the label ``bih'', thanks to the
relatively large tolerance of the encoding.

Accuracy in word learning increases when the perceived distance between
the labels is large relative to the tolerance of their encoding. This
improvement can occur in the same developmental stage if the perceptual
distance between labels is enhanced through using different-sounding
labels (e.g., ``lif'' vs. ``neem'' instead of ``bih'' and ``dih'') or
through using additional cues (Rost \& McMurray, 2009; Thiessen, 2007).
Besides, accuracy can improve across development through the refinement
of the encoding.

Building on this intuition, the current work proposes a precise
probabilistic model, which we use to both account for previous
experimental findings, and to make new predictions that have not been
tested before. Using new data collected from both preschool children and
adults, we show that the model can explain various patterns of learning
both within the same age and across development.

@@

\caption{\label{fig:illus} An illustration of the probabilistic/continuous account using simulated data. Two words are represented as distributions over the perceptual space. Here accuracy expresses the extent to which a given sound instance indicates the word on the left (red bar), as opposed to the word on the right (green bar). The values vary between 0.5 (total overlap between the distributions) and 1 (no overlap). Accuracy is low when the perceptual distance between labels is small relative to the variance. Accuracy increases when the perceptual distance is enhanced (through disambiguation), or when the variance decreases (e.g., through development).}\label{fig:illus}
\end{figure}
\end{CodeChunk}

@@

\subsection{Task}\label{task}

We model the word learning task introduced by Stager \& Werker (1997),
and a testing method similar to the one used by Yoshida et al. (2009).
In this task, participants are first exposed to the association between
pairs of nonsense words (e.g., ``lif''/``neem'') and pairs of objects.
After this exposure phase, participants perform a series of
two-alternative forced choices. In each testing trial, one of the two
sounds is uttered (e.g., ``lif'') and participants choose the
corresponding object from the two alternatives. An overview of the task
is shown in Figure \ref{fig:task}.

@@

\subsection{Simulations}\label{simulations}

During the testing phase (see Figure \ref{fig:task}), participants are
presented with one target sound \(s_T \in\{s_1, s_2\}\) and with both
objects \(o_1\) and \(o_2\). In order to make a choice, we need to
determine which object is more probable under the target sound \(s_T\).
In other words, we need to compare the probabilities \(P(o_1|s_T)\) and
\(P(o_2|s_T)\). The values of these probabilities can be computed by
summing over all possible concepts and labels:
\[P(o|s)=\sum_{C,L} P(o, C, L| s) \propto \sum_{C,L} P(o, C, L, s) \]
The joint probability \(P(o, C, L, s)\) is obtained by factoring the
Bayesian network in Figure \ref{fig:model}:
\[P(o,C,L,s) = P(s|L)P(L|C)P(C|o)P(o) \] which could be transformed
using Bayes rule into:

\[
P(o|s) = \frac{\sum_{C,L} P(s|L)P(o|C)P(L|C)}{\sum_{o} \sum_{C,L} P(s|L)P(o|C)P(L|C)}
\]

From this general expression we derive the exact analytical formula,
which expresses the probability of accurate responses in the testing
phase of the task:

\begin{equation}
P(o_T|s_T)= \frac{1 + e^{-(\Delta s^2 + \Delta o^2) /2\sigma^2}}{1 + e^{-(\Delta s^2 + \Delta o^2) /2\sigma^2} + e^{-\Delta s^2 /2\sigma^2} + e^{-\Delta o^2 /2\sigma^2 }}
\end{equation}

{\centering \includegraphics{figs/model-1}

\caption{\label{fig:model}Graphical representation of our model. Circles indicate random variables (shading indicates observed variables). The squares indicates fixed model parameters.}\label{fig:model}
\end{figure}
\end{CodeChunk}

In order to have a more quantitative understanding of the model, we
simulate the values of the predicted accuracy (Expression 1) as a
function of the perceptual distance between the sounds \(\Delta s\). We
used as parameters the two remaining variables, i.e., the visual
distance between the semantic referents \(\Delta o\) and the standard
deviation of the distributions \(p(s| L)\) and \(p(o | C)\) (which, in
this simulation, are assumed to have similar values, i.e.,
\(\sigma =\sigma_C \approx \sigma_L\)). The simulations are shown in
Figure \ref{fig:simulation}.

@@

\caption{\label{fig:simulation}The predicted probability of accurate responses in the testing phase, as a function the perceptual distance between the sounds. Colors indicate different values of the standard deviation which we assume is common to both the labels' and concepts' probabilistic representations. Panels represent graphs using different values of the visual distance between the objects.}\label{fig:simulation}
\end{figure*}
\end{CodeChunk}

The simulations explain some previousely documented facts, and make new
predictions:

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\item
  For fixed values of \(\Delta o\) and \(\sigma\), the probability of
  accurate responses increases as a function of \(\Delta s\). This
  pattern accounts for the fact that similar sounds are generally more
  challenging to learn than different sounds for both children (Stager
  \& Werker, 1997) and adults (Pajak et al., 2016).
\item
  For fixed values of \(\Delta s\) and \(\Delta o\), accuracy increases
  when the representational uncertainty (characterized with \(\sigma\))
  decreases. This fact may explain development, i.e., younger children
  have noisier representations (see Swingley, 2007; Yoshida et al.,
  2009), which leads to lower word recognition accuracy, especially for
  similar-sounding words.
\item
  For fixed values of \(\Delta s\) and \(\sigma\), accuracy increases
  with the visual distance between the semantic referents \(\Delta o\).
  This is a new prediction that our model makes. Previous work studied
  the effect of several bottom-up and top-down properties in
  disambiguating similar sounding words (Dautriche, Swingley, \&
  Christophe, 2015; Fennell \& Waxman, 2010; Rost \& McMurray, 2009;
  Thiessen, 2007), but no previous study tested the effect of the visual
  distance between the semantic referents.
\end{enumerate}

To sum, we introduced a model that accounts for some qualitative
learning patterns observed in previous studies, and makes a new
prediction. In the experiment below, we test whether the model makes
accurate \emph{quantitative} predictions by fitting it to new
experimental data collected from preschool children and adults.

@@

In this experiment, we tested participants in the word learning task
introduced above (Figure \ref{fig:task}). We explored all three
parameters of the model. In particular, both the sound similarity
(\(\Delta s\)) and object similarity (\(\Delta o\)) were varied
simultaneously in a within-subject design. Besides, two age groups
(preschool children and adults) were tested on the same task to explore
whether development can be characterized with the uncertainty parameter,
\(\sigma\), in the probabilsitic representations.

@@

\caption{\label{fig:stim}Normalized distances for both sound and object pairs used in this study.}\label{fig:stim}
\end{figure}
\end{CodeChunk}

@@

\caption{\label{fig:data_all}Accuracy of novel word recognition as as a function of the sound distance, the object distance, and the age group (preschool children vs. Adults). Experimental results are shown on the left. Predictions from Model 1 (one free parameter) and Model 2 (two free parameters) are shown in the middle and on the right, respectively.}\label{fig:all_data}
\end{figure*}
\end{CodeChunk}

@@

Preschoolers were asked if they would be willing to play a game on a
tablet with the experimenter and were informed that they could stop
playing at any time. The experimenter explained that the game consisted
in learning some words spoken in an alien planet. The experiment began
with two simple examples (not included in the analysis), and in these
examples children were given feedback from the experimenter so as to
make sure they correctly understood the structure of the task. After the
introduction and examples, children were tested in a sequence of five
conditions: the four experimental conditions plus the catch condition.
In each condition, participants saw a first block of four exposure
trials followed by four testing trials, and a second block of two
exposure trials (for memory refreshment) followed by an additional four
testing trials.

In the exposure trials, children saw two objects associated with their
corresponding sounds. We presented the first object on the left side of
the tablet's screen simultaneously with the corresponding sound. The
second sound-object association followed on the other side of the screen
after 500ms. For both objects, visual stimuli were present for the
duration of the sound clip (∼800ms). In the testing trials, children saw
both objects simultaneously and heard only one sound. They completed the
trial by selecting which of the two objects corresponded to the sound.
They responded by touching one of the pictures on the tablet.

The object-sound pairings were randomized across participants, as was
the order of the conditions (except for the catch condition which was
always placed in the middle of the testing sequence). We also randomized
the on-screen position (left vs.~right) of the two pictures on each
testing trial.

The procedure in the adult experiment was identical to that of preschool
children, except that preschoolers were accompanied by an experimenter
and used a tablet, whereas adults used their local computers to complete
the experiment online.

\subsubsection{Model fitting}\label{model-fitting}

We fit the probability function (equation 2) to the participants'
responses in each age group. The values of \(\Delta s\) and \(\Delta o\)
were set based on data from the similarity judgment task (Figure
\ref{fig:stim}). We used two models: \textbf{model 1} fit only one
parameter (\(\sigma = \sigma_C =\sigma_L\)), and \textbf{model 2} fit
two parameters (\(\sigma_C \neq \sigma_L\)). The values of the
parameters were derived using weighted least-squares estimates.

@@

We first analyzed the experimental results shown in Figure
(\ref{fig:data_all}, left), using a mixed-effects logistic regression
with sound and object distances as fixed effects, and with a maximal
random effects structure (Barr, Levy, Scheepers, \& Tily, 2013). Results
are shown in Table 1. We found a main effect of sound distance on the
accuracy of learning in both children and adults, thus replicating
previous findings. We also found a main effect of object distance, thus
confirming the new prediction of our model.

\begin{table}[!htbp] \centering 
  \caption{Estimates of predictor coefficients (and their standard errors) by age group in the regression model} 
  \label{} 
\begin{tabular}{@{\extracolsep{5pt}}lcc} 
\\[-1.8ex]\hline \\[-1.8ex] 
 & Children & Adults \\ 
\hline \\[-1.8ex] 
 (Intercept) & 0.426$^{*}$ (0.199) & 3.114$^{**}$ (1.015) \\ 
  Sound  & 0.272$^{**}$ (0.100) & 2.320$^{*}$ (0.981) \\ 
  Object & 0.315$^{*}$ (0.137) & 2.133$^{*}$ (0.952) \\ 
  Sound x Object & 0.151 (0.097) & 1.821 (0.976) \\ 
 \hline \\[-1.8ex] 
\textit{Note:}  & \multicolumn{2}{r}{$^{*}$p$<$0.05; $^{**}$p$<$0.01; $^{***}$p$<$0.001} \\ 
\end{tabular} 
\end{table}

Figures (\ref{fig:data_all}, middle and right graphs) show the
predictions of the models. Both model 1 and model 2 fit reasonably well
the experimental data in both children and adults. In particular, they
both correctly predict the relative recognition accuracy across
conditions: the pair of words that differ on both the object and sound
levels were the easiest to learn, followed by the pairs of words that
differ on only one level, then the pair of words that are similar on
both levels. Note, however, that both models predict a slight
interaction between sounds and objects, but such an interaction was not
significant in the the experimental data (Table 1).

For Model 1, the fitted uncertainty parameter was \(\sigma =\) 0.63
{[}0.53, 0.73{]}\footnote{95 \% Confidence intervals.} for preschoolers,
and \(\sigma =\) 0.16 {[}0.12, 0.19{]} for adults. As predicted, the
uncertainty parameter was larger for children than it was for adults.
The model explained the majority of the variance (\(R^2=\) 0.94). For
model 2, children had a label-specific uncertainty of \(\sigma_S =\) 0.9
{[}0.68, 1.11{]}, and a concept-specific uncertainty of \(\sigma_C =\)
0.29 {[}0.1, 0.49{]}. Adults had a label-specific noise of
\(\sigma_S =\) 0.16 {[}0.05, 0.28{]}, and a concept-specific noise of
\(\sigma_C =\) 0.14 {[}0.05, 0.23{]}. The model explained almost all the
variance (\(R^2=\) 0.96). Interestingly, despite the fact that Model 1
used only one degree of freedom, it captures the data variance almost as
well as Model 2 (which used two degrees of freedom).

@@

Since the seminal work of Stager \& Werker (1997) showing a discrepancy
between children's speech perception abilities and their word learning
skills, researchers have debated about the source of such a discrepancy.
In this work, we were more interested in the development of this
phenomenon, regardless of its exact origin. Building on some previous
findings and discussions (Swingley, 2007; Yoshida et al., 2009), we
proposed a model which assumes that subtle perceptual features are
correctly encoded during word meaning learning, but that the precision
of this encoding vary across development. We compared the predictions to
new data which we collected from preschool children and adults. We found
that the probabilistic model captured not only the patterns of
development, but also various learning patterns within the same age
group.

The model made a new prediction which we tested experimentally for the
first time: learning similar words is not only modulated by to
similarity of their phonological forms, but also by the visual
similarity of their semantic referents. More generally, since visual
similarity is an early organizing feature in the semantic domain (e.g.,
Wojcik \& Saffran, 2013), our finding suggests that children may
prioritize the acquisition of words that are quite distant in the
semantic space. This suggestion is supported by recent findings based on
the investigation of early vocabulary growth (Engelthaler \& Hills,
2017; Sizemore, Karuza, Giusti, \& Bassett, 2018). That said, further
work is needed to explore the effect on word learning of other semantic
dimensions that could be encoded by children (e.g.,
conceptual/functional features).

To conclude, this paper proposes a computational model for the
probabilistic/continuous account, and the findings do endorse this
account as a valid framework for the study of similar word learning and
the mechanism of its development. More generally, together with previous
modeling efforts (e.g., Apfelbaum \& McMurray, 2011; Hofer \& Levy,
2017), our work highlight the crucial role computation models can play
in theory building through both a compact organization of existing
knowledge and through generating new principled hypothesis.

\vspace{1em}

\fbox{\parbox[b][][c]{7.3cm}{\centering All data and code for these analyses are available at\ \url{https://github.com/afourtassi/networks}}}
\vspace{1em}

\section{Acknowledgements}\label{acknowledgements}

This work was supported by a post-doctoral grant from the Fyssen
Foundation, NSF \#1528526, and NSF \#1659585.

@@

\hypertarget{refs}{}
\hypertarget{ref-apfelbaum2011}{}
Apfelbaum, K., \& McMurray, B. (2011). Using variability to guide
dimensional weighting: Associative mechanisms in early word learning.
\emph{Cognitive Science}, \emph{35}(6).

\hypertarget{ref-barr2013}{}
Barr, D., Levy, R., Scheepers, C., \& Tily, H. (2013). Random effects
structure for confirmatory hypothesis testing: Keep it maximal.
\emph{Journal of Memory and Language}, \emph{68}(3).

\hypertarget{ref-dautriche2015}{}
Dautriche, I., Swingley, D., \& Christophe, A. (2015). Learning novel
phonological neighbors: Syntactic category matters. \emph{Cognition},
\emph{143}.

@@

>>>>

\title{The role of continuous developmental change in explaining
discontinuities in word learning}

\usepackage{tipa}
\usepackage[rightcaption]{sidecap}

@@

\begin{abstract}
Cognitive development is often characterized in term of discontinuities.
Nevertheless, in some cases, this characterization can be problematic
and unnecessary. In this paper, we show that a phenomenon that appears
stage-like, can be well described by a continuous account. We used as a
case study the finding by Stager and Werker (1997) that children's early
ability to distinguish similar sounds does not automatically translate
into word learning skills. Early explanations proposed that children may
not be able to encode subtle phonetic contrasts when learning novel word
meanings, thus suggesting a discontinuous/stage-like pattern of
development. However, later work has revealed (e.g., through using
simpler testing methods) that children do encode such contrasts, thus
favoring a continuous pattern of development. Here we propose a precise
probabilistic model describing how development may proceed in a
continuous fashion across the lifespan. The model accounts for
previously documented facts and provides new predictions. We collected
data from preschool children and adults, and we showed that the model
can explain various patterns of learning both within the same age and
across development. The findings suggest that other aspects of cognitive
development which are typically thought of as discontinuities, may
emerge from simpler, continuous mechanisms.

@@

Cognitive development is sometimes characterized in terms of a
succession of discontinuous stages (Piaget, 1954). Such characterization
may be problematic as it often renders the study of the learning
mechanism quite challenging. Indeed, it is unclear how a precise
algorithm----whose steps are gradual and logically related----can
account for transitions between stages involving qualitatively different
representations (e.g., Carey, 2009).

Nevertheless, at least in some cases, development may only appear to be
stage-like. This appearance can be due, for example, to the use of a
cognitively-demanding task which may mask learning, or to the use of
hard statistical thresholding (in particular, p-value \textless{} 0.05)
which creates a spurious dichotomy between success and failure in
observing a given behavior. In such cases, positing discontinuous stages
is unnecessary. Instead, as we shall see in this paper, a continuous
model----involving similar representations across the lifespan----may
provide a simpler and more transparent account of development.

We use a case study from word learning literature. Stager \& Werker
(1997) first showed that children's early ability to distinguish similar
sounds (``dih'' vs. ``bih'') does not automatically translate into word
learning skills (i.e., associating these sounds with different
meanings). Subsequent research has focused on proposing possible
explanations for this observed gap between speech perception and word
learning (e.g., Fennell \& Waxman, 2010; Hofer \& Levy, 2017; Rost \&
McMurray, 2009; Stager \& Werker, 1997).

In this work we focus, rather, on the development of this phenomenon. In
fact, regardless of the exact explanation, it is generally accepted that
by around 17 m.o, children succeed in the same task (Werker, Fennell,
Corcoran, \& Stager, 2002). How does development proceed? Early accounts
assumed that children encode words in a binary way: they either fail or
succeed in encoding the relevant phonetic details (simultaneously with
the meanings). This account suggested a discontinuous/stage-like pattern
of development whereby younger children fail to encode the contrastive
phonetic detail, whereas older children succeed.

Nevertheless, subsequent findings have suggested otherwise. On the one
hand, 14-month-olds--who typically fail in the original task--succeed
when an easier testing method is used, even under the same learning
conditions (Yoshida, Fennell, Swingley, \& Werker, 2009). They also
succeed when uncertainty is mitigated via disambiguating cues (e.g.,
Thiessen, 2007). On the other hand, adults show patterns of learning
similar to those shown by 14-month-olds when the task is more
challenging and when similarity between words increases (Pajak, Creel,
\& Levy, 2016; White, Yee, Blumstein, \& Morgan, 2013).

These evidence point towards another scenario, where the representations
are encoded in a probabilistic (rather than binary) way, and where
development is continuous, rather than stage-like (Pajak et al., 2016;
Swingley, 2007; Yoshida et al., 2009). On this account, correct
representations are learned early in development, but these
representations are encoded with higher uncertainty in younger children,
leading to apparent failure in relatively demanding tasks. Development
is a continuous process whereby the initial noisy representations become
more precise. Crucially, more precise representation are still
imperfect: Even adults show low accuracy learning when the sounds are
subtle enough, e.g., non-native sounds (Pajak et al., 2016).

We provide an intuitive illustration of how such an account explains
patterns of learning and development in Figure \ref{fig:illus}. We
observe low accuracy in word learning when the perceptual distance
between the labels is small relative to the uncertainty with which these
labels are encoded. For example, in Stager and Werker's original
experiment, children are supposed to associate label 1 (``bih'') and
label 2 (``dih'') with object 1 and object 2, respectively. Though
children could learn that the label ``bih'' is a better match to object
1 than ``dih'', they could still judge the sound ``dih'' as a plausible
instance of the label ``bih'', thanks to the relatively large
uncertainty of the encoding.

According to this account, accuracy in word learning improves if we
increase either the perceptual distinctiveness of the stimuli (e.g.,
through using different-sounding labels), or the precision of the
encoding itself (e.g., across development). Building on this intuition,
the current work proposes a precise probabilistic model, which we use to
both account for previous experimental findings, and to make new
predictions that have not been tested before. Using new data collected
from both preschool children and adults, we show that the model can
explain various patterns of learning both within the same age and across
development.

@@

\caption{\label{fig:illus} An illustration of the probabilistic/continuous account using simulated data. A word is represented with a distribution over the perceptual space (indicated in red or green). When the uncertainty of the representation is large relative to the distance between the stimuli (top panel), an instance of the red category (indicated with a star) could also be a plausible instance of the green category, hence the low recognition accuracy score. The accuracy increases when the stimuli are less similar (left panel), or when the representation are more precise (right panel).}\label{fig:illus}
\end{figure}
\end{CodeChunk}

@@

\begin{SCfigure}
\centering
\includegraphics[width=1.5in]{figs/model.png}
\caption{Graphical representation of our model. Circles indicate random variables (shading indicates observed variables). The squares indicates fixed model parameters.}
\label{fig:model}
\end{SCfigure}

\subsection{Inference}\label{inference}

The learner hears a sound \(s\) and has to decided which object \(o\)
provides an optimal match to this sound. To this end, the learner has to
compute the probability \(P(o|s)\) for all possible objects. This
probability can be computed by summing over all possible concepts and
labels:
\[P(o|s)=\sum_{C,L} P(o, C, L| s) \propto \sum_{C,L} P(o, C, L, s) \]
The joint probability \(P(o, C, L, s_T)\) is obtained by factoring the
Bayesian network in Figure \ref{fig:model}:
\[P(o,C,L,s) = P(s|L)P(L|C)P(C|o)P(o) \] which could be transformed
using Bayes rule into:

\begin{equation}
P(o|s) = \frac{\sum_{C,L} P(s|L)P(o|C)P(L|C)}{\sum_{o} \sum_{C,L} P(s|L)P(o|C)P(L|C)}
\end{equation}

\subsection{Task and model
predictions}\label{task-and-model-predictions}

@@

We use the model to predict performance in the word learning task
introduced by Stager \& Werker (1997), and a testing method similar to
the one used by Yoshida et al. (2009). In this task, participants are
first exposed to the association between pairs of nonsense words (e.g.,
``lif''/``neem'') and pairs of objects. After this exposure phase,
participants perform a series of test trials. In each of these trials,
one of the two sounds is uttered (e.g., ``lif'') and participants choose
the corresponding object from the two alternatives. An overview of the
task is shown in Figure \ref{fig:task}.

We use the general Equation 1 to derive the exact analytical expression
for the probability of accurate responses \(p(o_T | s)\) (target object
\(o_T\) given a sound \(s\)) in the simple case of two-alternative
forced choice in the testing phase of our experimental task:

\begin{equation}
P(o_T|s)= \frac{1 + e^{-(\Delta s^2/2\sigma_L^2+ \Delta o^2/2\sigma_C^2)}}{1 + e^{-(\Delta s^2/2\sigma_L^2+ \Delta o^2/2\sigma_C^2)}+ e^{-\Delta s^2 /2\sigma_L^2} + e^{-\Delta o^2 /2\sigma_C^2 }}
\end{equation}

Figure \ref{fig:simulation} show simulations of the predicted accuracy
(Expression 2) as a function of the distinctiveness parameters
(\(\Delta s\) and \(\Delta o\)) and the precision parameter, i.e., the
variance of the distributions \(p(s| L)\) and \(p(o | C)\) (which, for
ease of the simulations' interpretation, are assumed to have similar
values, i.e., \(\sigma =\sigma_C \approx \sigma_L\)).

@@

\caption{\label{fig:simulation}The predicted probability of accurate responses in the testing phase as a function of stimuli distinctiveness $\Delta s$ (x-axis) and $\Delta o$ (different panels) and representation precision $\sigma$ (color). Dashed line represents chance.}\label{fig:simulation}
\end{figure*}
\end{CodeChunk}

The simulations explain some previously documented facts, and make new
predictions:

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\item
  For fixed values of \(\Delta o\) and \(\sigma\), the probability of
  accurate responses increases as a function of \(\Delta s\). This
  pattern accounts for the fact that similar sounds are generally more
  challenging to learn than different sounds for both children (Stager
  \& Werker, 1997) and adults (Pajak et al., 2016).
\item
  For fixed values of \(\Delta s\) and \(\Delta o\), accuracy increases
  when the representational uncertainty (characterized with \(\sigma\))
  decreases. This fact may explain development, i.e., younger children
  have noisier representations (see Swingley, 2007; Yoshida et al.,
  2009), which leads to lower word recognition accuracy, especially for
  similar-sounding words.
\item
  For fixed values of \(\Delta s\) and \(\sigma\), accuracy increases
  with the visual distance between the semantic referents \(\Delta o\).
  This is a new prediction that our model makes. Previous work studied
  the effect of several bottom-up and top-down properties in
  disambiguating similar sounding words (e.g., Fennell \& Waxman, 2010;
  Rost \& McMurray, 2009; Thiessen, 2007), but no previous study tested
  the effect of the visual distance between the semantic referents.
\end{enumerate}

In sum, we introduced a model that accounts for some qualitative
learning patterns observed in previous studies, and makes a new
prediction. In the experiment below, we test whether the model makes
accurate quantitative predictions by fitting it to new experimental data
collected from preschool children and adults.

@@

In this experiment, we tested participants in the word learning task
introduced above (Figure \ref{fig:task}). We explored all three
parameters of the model. In particular, both the sound similarity
(\(\Delta s\)) and object similarity (\(\Delta o\)) were varied
simultaneously in a within-subject design. Two age groups (preschool
children and adults) were tested on the same task to explore whether
development can be characterized with the uncertainty parameter,
\(\sigma\), in the probabilistic representations. The Experiment and the
predictions were pre-registered.

@@

\caption{\label{fig:stim}Distances for both sound and object pairs from an adult norming study. Error bars represent 95\% confidence intervals.}\label{fig:stim}
\end{figure}
\end{CodeChunk}

@@

\caption{\label{fig:data_all}Accuracy of novel word recognition as as a function of the sound distance, the object distance, and the age group (preschool children vs. Adults). We show both experimental results (top) and model predictions (bottom). Error bars represent 95\% confidence intervals.}\label{fig:all_data}
\end{figure*}
\end{CodeChunk}

@@

Preschoolers were tested at the nursery school using a tablet, whereas
adults used their local computers to complete the same experiment
online. Participants were tested in a sequence of five conditions: the
four experimental conditions plus the catch condition. In each
condition, participants saw a first block of four exposure trials
followed by four testing trials, and a second block of two exposure
trials (for memory refreshment) followed by an additional four testing
trials.

In the exposure trials, participants saw two objects associated with
their corresponding sounds. We presented the first object on the left
side of the tablet's screen simultaneously with the corresponding sound.
The second sound-object association followed on the other side of the
screen after 500ms. For both objects, visual stimuli were present for
the duration of the sound clip (∼800ms). In the testing trials,
participants saw both objects simultaneously and heard only one sound.
They completed the trial by selecting which of the two objects
corresponded to the sound. The object-sound pairings were randomized
across participants, as was the order of the conditions (except for the
catch condition which was always placed in the middle of the testing
sequence). We also randomized the on-screen position (left vs.~right) of
the two pictures on each testing trial.

@@

We first analyzed the experimental results shown in Figure
(\ref{fig:data_all}, top), using a mixed-effects logistic regression
with sound distance, object distance and age group as fixed effects, and
with a maximal random effects structure (allowing us to take into
account the full nested structure of our data) (Barr, Levy, Scheepers,
\& Tily, 2013). We found main effects for all the fixed effects, i.e.,
the sound distance (\(\beta =\) 0.68, \(p\) \textless{}
0.001)---replicating previous findings, the object distance (\(\beta =\)
0.6, \(p\) \textless{} 0.001)---confirming the new prediction of our
model, and age group (\(\beta =\) 0.59, \(p\) \textless{}
0.001)---showing that performance improves with age. Besides, we found
two-way interactions between sound and object distances (\(\beta =\)
0.36, \(p\) \textless{}0.05) and between sound distance and age
(\(\beta =\) 0.37, \(p\) \textless{}0.01).

Second, we fit the probability function (Equation 2) to the
participants' responses in each age group. The values of \(\Delta s\)
and \(\Delta o\) were set based on data from the similarity judgment
task (Figure \ref{fig:stim}). Thus, the model has two degrees of freedom
in each group, i.e., \(\sigma_C\) and \(\sigma_L\). Figure
\ref{fig:data_all} (bottom) show the predictions of the model. The model
correctly predicts the relative recognition accuracy across conditions:
the pair of words that differ on both the object and sound levels were
the easiest to learn, followed by the pairs of words that differ on only
one level, then the pair of words that are similar on both levels. At
the quantitative level, the models explained almost all the variance in
the participants' mean responses (\(R^2=\) 0.96).

We also fit a simplified model with only one degree of freedom
(\(\sigma = \sigma_C =\sigma_L\)). This second model is not intended to
test any theoretical hypothesis. We use it, rather, as a baseline to
understand how much of the explanatory power of the model is due to the
extra degree of freedom. This baseline model still captured the
qualitative pattern (graph not shown), and explained the majority of the
variance (\(R^2=\) 0.94). Besides, model comparison using Akaike
information criterion yielded \(AIC(model) - AIC(baseline) =\) -3.79 for
children data and \(AIC(model) - AIC(baseline) =\) 1.62 for adult data.
These numbers suggest that the model remains highly predictive even with
one degree of freedom.

As for the values of the parameters, children had a label-specific
uncertainty of \(\sigma_S =\) 0.9 {[}0.68,
1.11{]}\footnote{All uncertainty intervals in this paper represent 95\% Confidence Intervals.},
and a concept-specific uncertainty of \(\sigma_C =\) 0.29 {[}0.1,
0.49{]}. Adults had a label-specific uncertainty of \(\sigma_S =\) 0.16
{[}0.05, 0.28{]}, and a concept-specific uncertainty of \(\sigma_C =\)
0.14 {[}0.05, 0.23{]}. As predicted, the uncertainty parameters were
larger for children than they were for adults.

@@

This paper studies the hypothesis that some seemingly stage-like
patterns in cognitive development can be characterized in a continuous
fashion. We used as a case study the seminal work of Stager \& Werker
(1997) showing a discrepancy between children's speech perception
abilities and their word learning skills. While most previous work
debated the source of this discrepancy, here we were more interested in
its mechanism of development.

Building on some previous findings and discussions (e.g., Swingley,
2007; Yoshida et al., 2009), we proposed a model which assumes that
subtle perceptual features are correctly encoded during word meaning
learning, but that the precision of this encoding improves across
development. We tested the model's predictions against data collected
from preschool children and adults and we showed that development can be
characterized as a continuous refinement in qualitatively similar
representations across the life span.

The model made a new prediction which we tested experimentally for the
first time: learning similar words is not only modulated by the
similarity of their phonological forms, but also by the visual
similarity of their semantic referents. More generally, since visual
similarity is an early organizing feature in the semantic domain (e.g.,
Wojcik \& Saffran, 2013), our finding suggests that children may
prioritize the acquisition of words that are quite distant in the
semantic space. This suggestion is supported by recent findings based on
the investigation of early vocabulary growth (Engelthaler \& Hills,
2017; Sizemore, Karuza, Giusti, \& Bassett, 2018). That said, further
work is needed to explore the effect on word learning of other semantic
dimensions that could be encoded by children (e.g.,
conceptual/functional features).

One limitation of this work is that the model was fit to data from
children at a relatively older age (4-5 years old) than what is
typically studied in the literature (around 14 month-old). While an
important conclusion of this paper is that the same process may apply to
different developmental stages, it would be interesting to
experimentally verify the generality of this conclusion in the context
of early language acquisition. A challenge to this next step would be to
construct a computational framework which would compare performance
across heterogeneous testing methods (e.g., looking time
vs.~forced-choice).

To sum, this paper proposes a model that accounts for the development of
an important aspect of word learning. The findings show that data can be
explained based on a continuous process operating over similar
representations across development. We used a case from word learning as
an example, but the same idea might apply to other aspects of cognitive
development which are tyipcally thought of as stage-like (e.g.,
acquisition of a theory of mind). A crucial task would be to investigate
the extent to which such discontinuities emerge due to genuine
qualitative changes, and the extent to which they reflect the
granularity of the researchers' own measurement tools.

@@

\hypertarget{refs}{}
\hypertarget{ref-barr2013}{}
Barr, D., Levy, R., Scheepers, C., \& Tily, H. (2013). Random effects
structure for confirmatory hypothesis testing: Keep it maximal.
\emph{Journal of Memory and Language}, \emph{68}(3).

\hypertarget{ref-carey2009}{}
Carey, S. (2009). \emph{The origin of concepts}. Oxford University
Press.

@@

\hypertarget{ref-piaget1954}{}
Piaget, J. (1954). \emph{The construction of reality in the child}. New
York, NY, US: Basic Books.
